Concurrent, Parallel Garbage Collection in Linear Time

### 2016-06-02 finaletd/prelim.tex
prelim changes

Distributed systems require garbage collectors to reclaim memory. Distributed Garbage Collector(DGC) requires an algorithm to be designed in a different way than the other approaches. Most of the algorithm discussed above cannot be applicable to the distributed setting. Bevan\cite{Bevan87} proposed the use of Weighted Reference Counting as an efficient solution to DGC. Each reference has two weights: a partial weight and a total weight. Every reference creation halves the partial weight. The drawback of this method is an application cannot have more than 32 or 64 references based on the architecture. Piquer\cite{piquer91} suggests an original solution to the previous approach. Both of the above-mentioned methods use straight reference counting, so they cannot reclaim cyclic structures and they are not resilient to message failures. Shapiro et al \cite{Shapiro92} provided an acyclic truly distributed garbage collector. Although this method is not cyclic distributed garbage collector, the work lays a foundation for a model of distributed garbage collection. A standard approach to a distributed tracing collector is to combine independent local, per-space collectors, with a global inter-space collector. The main problem with distributed tracing is to synchronize the distributed mark phase with independent sweep phase\cite{plain95}. Hughes designed a distributed tracing garbage collector with a time-stamp instead of mark bits\cite{hugh85}. The method requires the global GC to process nodes only when the application stops its execution or requests a stop when collecting. This is not a practical solution for many systems. Several solutions based on a centralized collector and moving objects to one space are proposed\cite{Maheshwari1997,Maheshwari1997b,Liskov,Ladin,Veiga05}. Recently, a mobile actor-based distributed garbage collector was designed\cite{Wang2006}. Mobile actor DGC cannot be easily implemented as an actor need access to remote site root set and actors move with huge memory.  All of the distributed garbage collection solutions have one or more of the problems: inability to detect cycle, moving objects to one site for cycle detection, weak heuristics to compute cycle, no fault-tolerance, centralized solution, stopping the world, not scalable, and failure to be resilient to message failures.

\section{Cycle detection using Strong-Weak}
Brownbridge classifies the edges in the reference graph into strong and weak. The classification is based on the invariant that there is no cycle of strong edges. So every cycle must contain at least one weak edge. The notion of the strong edge is connected to liveness of a node. So  the existence of strong incoming edge for a node indicates that the node is live.

Strong edges form a connected directed acyclic graph in which every node is reachable from R. The rest of the edges are classified as weak. This classification is not simple to compute as themselves might require complete scanning of the graph. To avoid complete scanning, some heuristic approaches are required. The identification of cycle creating edges are the bottleneck of the classification problem. The edges are  labeled weak if it is created to a node that already exists in the memory. This heuristic guarantee that all cycle contains at least a weak edge. But a node with weak incoming edge does not mean it is part of a cycle. The heuristic helps the mutator to save time in classifying the edges. But the collectors require more information about the topology of the subgraph to identify if the subgraph is indeed garbage to delete.

The heuristic idea mentioned above solves the problem and graph will not contain a cycle of strong edges at all point in time. The other technique required by Brownbridge algorithm is the ability to change the strong and weak edges whenever required by the nodes. To solve this issue, Brownbridge uses two boolean values one for each edge in the source of the edge and one for each node. For an edge to strong, both the boolean value has to be same. To change strong to weak, the boolean values must differ. So node that wants to change the type of the edge, a node changes the corresponding boolean value. A node will change all the incoming edge type when it changes the boolean value of it. A node by changing the boolean value for outgoing edge changes the type of only the edge whose boolean value is changed. The idea of toggling between the edge type is an important piece of the Brownbridge's algorithm.

>>>>

Distributed systems require garbage collectors to reclaim memory. Distributed Garbage Collector(DGC) requires an algorithm to be designed in a different way than the other approaches. Most of the algorithm discussed above cannot be applicable to the distributed setting. Bevan\cite{Bevan87} proposed the use of Weighted Reference Counting as an efficient solution to DGC. Each reference has two weights: a partial weight and a total weight. Every reference creation halves the partial weight. The drawback of this method is an application cannot have more than 32 or 64 references based on the architecture. Piquer\cite{piquer91} suggests an original solution to the previous approach. Both of the above-mentioned methods use straight reference counting, so they cannot reclaim cyclic structures and they are not resilient to message failures. Shapiro et al \cite{Shapiro92} provided an acyclic truly distributed garbage collector. Although this method is not cyclic distributed garbage collector, the work lays a foundation for a model of distributed garbage collection. A standard approach to a distributed tracing collector is to combine independent local, per-space collectors, with a global inter-space collector. The main problem with distributed tracing is to synchronize the distributed mark phase with independent sweep phase\cite{plain95}. Hughes designed a distributed tracing garbage collector with a time-stamp instead of mark bits\cite{hugh85}. The method requires the global GC to process nodes only when the application stops its execution or requests a stop when collecting. This is not a practical solution for many systems. Several solutions based on a centralized collector and moving objects to one space are proposed\cite{Maheshwari1997,Maheshwari1997b,Liskov,Ladin,Veiga05}. Recently, a mobile actor-based distributed garbage collector was designed\cite{Wang2006}. Mobile actor DGCs cannot be easily implemented as an actor needs access to a remote site's root set and actors move with huge memory payload.  All of the distributed garbage collection solutions have one or more of the problems: inability to detect cycles, moving objects to one site for cycle detection, weak heuristics to compute cycles, no fault-tolerance, centralized solution, stopping the world, not scalable, and failure to be resilient to message failures.

\section{Cycle detection using Strong-Weak}
Brownbridge classifies the edges in the reference graph into strong and weak. The classification is based on the invariant that there is no cycle of strong edges. So every cycle must contain at least one weak edge. The notion of the strong edge is connected to the liveness of the node. So  the existence of a strong incoming edge for a node indicates that the node is live.

Strong edges form a directed acyclic graph in which every node is reachable from R. The rest of the edges are classified as weak. This classification is not simple to compute as they themselves might require complete scanning of the graph. To avoid complete scanning, some heuristic approaches are required. The identification of cycle creating edges are the bottleneck of the classification problem. The edges are  labeled weak if it is created to a node that already exists in the memory. This heuristic guarantee that all cycles contain at least one weak edge. But a node with weak incoming edge does not mean it is part of a cycle. The heuristic helps the mutator to save time in classifying the edges. But the collectors require more information about the topology of the subgraph to identify whether the subgraph is indeed garbage.

The heuristic idea mentioned above prevents the creation of strong cycles. The other technique required by the Brownbridge algorithm is the ability to change the strong and weak edges whenever required by the nodes. To solve this issue, Brownbridge used two boolean values, one for each edge in the source and one for each node. For an edge to be strong, both the boolean values have to be same. So node that wants to change the type of the edge, a node changes the corresponding boolean value. A node will change all the incoming edge type when it changes the boolean value of it. A node by changing the boolean value for outgoing edge changes the type of only the edge whose boolean value is changed. The idea of toggling between the edge type is an important piece of the Brownbridge's algorithm.
